Modeling Two Body Collisions with Spring Potential Interactions

### 2014-12-12 paper/paper.tex
edited paper

In addition to the position and velocity of the masses we also calculate their kinetic energies and the potential energy stored in the deformed bodies, e.g. spring \eqref{eq:energy}. The force is conservative, so $\mathbf{F} = - \nabla V \rightarrow V = \int \left|F\right| d\left(\Delta x\right)$. With a small enough $\Delta t$ these iterative equations model the physics of the interaction very well. To ensure that $\Delta t$ is not set too large the code checks each masses change in position at every time step. If the change in position is larger than 1\% of the masses radius a warning is generated and the simulation is re-run with a smaller $\Delta t$
\begin{equation}\label{eq:energy}
  V_n = \frac{k}{\lambda+1}\left(\Delta x_n\right)^{\lambda+1},\quad T_n = \frac{m_1}{2}\left(\dot{\mathbf{x}}^{(1)}_n\right)^2 + \frac{m_2}{2}\left(\dot{\mathbf{x}}^{(2)}_n\right)^2
\end{equation}

\section{Results}
\subsection{Simulation Performance}
One of the first tasks was to validate the simulation framework using a system where the physics was simple and easy to intuitively predict. This system uses a stationary target of radius units 2, with a projectile of radius units 1 incoming at velocity units 5 with an impact parameter of 1 unit. The objects have equal density. We configure this system to have no energy loss ($\eta = 0$), $\lambda=1$ power law, and $k=100$ force constant. First we visualize the track points for each objects tracjectory while they are in contact with each other. We show this visualization in Figure~\ref{fig:tracks_default}.
\begin{figure}[h!]
  \includegraphics[width=.45\textwidth]{plots/default/x_vs_y_with_ellipse.pdf}
  \includegraphics[width=.45\textwidth]{plots/default/x_vs_y_with_ellipse_MS.pdf}
                  {\par\nobreak\rule[9pt]{35em}{0.5pt}\vspace{-5mm}}
                  \caption{(L) The initial state of the system when the objects come in contact. The collections of points (which form a continuous line) show the position of the center of each object at each time step while the objects are in contact. (R) The state of the system when the objects are at a minimum separation (maximum $\Delta x$ from equilibrium).}
                  \label{fig:tracks_default}
\end{figure}
We also check for kinematic trends to be physically sound for the given initial conditions and parameters of the system.
\begin{figure}
  \includegraphics[width=.8\textwidth]{plots/default/E_vs_t.pdf}
  {\par\nobreak\rule[9pt]{35em}{0.5pt}\vspace{-5mm}}
  \caption{Trends for different energies in the system: object and total kinetic energies, potential energy, and the total energy of the system.}
\end{figure}

>>>>

In addition to the position and velocity of the masses we also calculate their kinetic energies and the potential energy stored in the deformed bodies, e.g. spring \eqref{eq:energy}. The force is conservative, so $\mathbf{F} = - \nabla V \rightarrow V = \int \left|F\right| d\left(\Delta x\right)$. With a small enough $\Delta t$ these iterative equations model the physics of the interaction very well. To ensure that $\Delta t$ is not set too large the code checks each masses change in position at every time step. If the change in position is larger than 1\% of the masses radius a warning is generated and the simulation is re-run with a smaller $\Delta t$. Throughout the rest of this paper SI units are assumed where none are given.
\begin{equation}\label{eq:energy}
  V_n = \frac{k}{\lambda+1}\left(\Delta x_n\right)^{\lambda+1},\quad T_n = \frac{m_1}{2}\left(\dot{\mathbf{x}}^{(1)}_n\right)^2 + \frac{m_2}{2}\left(\dot{\mathbf{x}}^{(2)}_n\right)^2
\end{equation}

\section{Results}
\subsection{Simulation Performance}
\subsubsection{Default Configuration}
One of the first tasks was to validate the simulation framework using a system where the physics was simple and easy to intuitively predict. This system uses a stationary target of radius 2, with a projectile of radius 1 incoming at velocity 5 with an impact parameter of 1. The objects have equal density. We configure this system to have no energy loss ($\eta = 0$), $\lambda=1$ power law, and $k=100$ spring constant. To begin to get an understanding of the collision we first plot each objects center of mass tracjectory during the time when they are in contact with each other and are undergoing accelerations (Figure~\ref{fig:tracks_default}). The projectile rebounds off the target sending it off at a slow velocity in the oposite direction will the systems center of mass steadly moves in a straight line.
\begin{figure}[h!]
  \includegraphics[width=.45\textwidth]{plots/default/x_vs_y_with_ellipse.pdf}
  \includegraphics[width=.45\textwidth]{plots/default/x_vs_y_with_ellipse_MS.pdf}
                  {\par\nobreak\rule[9pt]{35em}{0.5pt}\vspace{-5mm}}
                  \caption{(L) The initial state of the system when the objects come in contact. The collections of points (which form a continuous line) show the position of the center of each object at each time step while the objects are in contact. (R) The state of the system when the objects are at a minimum separation i.e. maximum $\Delta x$ from equilibrium.}
                  \label{fig:tracks_default}
\end{figure}

discuss energy here

\begin{figure}
  \includegraphics[width=.8\textwidth]{plots/default/E_vs_t.pdf}
  {\par\nobreak\rule[9pt]{35em}{0.5pt}\vspace{-5mm}}
  \caption{Trends for different energies in the system: object and total kinetic energies, potential energy, and the total energy of the system.}
\end{figure}

\subsubsection{c6 Equal Masses, Zero Momentum}
the

\subsubsection{c7}
the
